Make any stupid plot you want

### 2020-12-08 draft/sections/intro.tex
clean up, moving around
rewrote intro to math section w/ more context
taus

\documentclass[letterpaper,onecolumn,titlepage]{Ythesis}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usetikzlibrary{cd}
\usepackage{array,multirow}
\usepackage{subcaption}
\usepackage{subfiles}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
\usepackage{diagbox}
\usepackage{graphicx}
\usepackage[backend=bibtex, style=numeric-comp]{biblatex}
\bibliography{glasslab_viz}
\title{Make any stupid plot you want}
\author{Hannah Aizenman}
\committee{Dr. Michael Grossberg (Advisor), Dr. Robert Haralick, Dr. Lev Manovich, Dr. Huy Vo}
\submitted{}
\abstract{}

\begin{document}
\makefrontmatter

\section{Introduction}
\label{sec:introduction}
\begin{figure}
    \includegraphics[width=.5\textwidth]{figures/intro/viz_same.png}
    \caption[]{Implicit in visualization is the assumption that these three representations of data are equivalent, specifically that the measurements within a variable and relations of the measurements of each variable are preserved. }
    \label{fig:viz_same}
\end{figure}

\subsection{Thesis statement}
We define a visualization as a transform from data to graphic that preserves the topology of the data and the properties of the measurement type. In fig~\ref{fig:viz_same}, we implicitly assume that the translation from table to heatmap has preserved the order of observations (the rows) and that the perceptually uniform sequential colormap has been applied such that the ordering relation on floats matches the ordering on the colormap (darker colors map to larger numbers). We also make this assumption about color in the scatter map, and that the translation to size and position on screen also respect the ordering on floats. In this work, we propose to mathematically describe the transform of data to visual space such that we can make explicit the implicit topology and types visualizations preserve. We then propose a new architecture for the Python visualization library Matplotlib \cite{hunterMatplotlib2DGraphics2007} based on these descriptions because the Matplolib artist layer is analogous to the transforms.

\subfile{sections/math}
\printbibliography
\end{document}

>>>>

\documentclass[../main.tex]{subfiles}
\begin{document}
\section{Introduction}
\label{sec:introduction}
\begin{figure}
    \includegraphics[width=.5\textwidth]{figures/intro/viz_same.png}
    \caption[]{Implicit in visualization is the assumption that these three representations of data are equivalent, specifically that the measurements within a variable and relations of the measurements of each variable are preserved. }
    \label{fig:viz_same}
    %% change up image to dubois or bertin or

\end{figure}

\subsection{Thesis statement}
We define a visualization as a transform from data to graphic that preserves the topology of the data and faithfully map the properties of the measurement type. In fig~\ref{fig:viz_same}, we implicitly assume that the translation from table to heatmap has preserved the order of observations (the rows) and that the perceptually uniform sequential colormap has been applied such that the ordering relation on floats matches the ordering on the colormap (darker colors map to larger numbers). We also make this assumption about color in the scatter map, and that the translation to size and position on screen also respect the ordering on floats. In this work, we propose to mathematically describe the transform of data to visual space such that we can make explicit the implicit topology and types visualizations preserve. We then propose a new architecture for the Python visualization library Matplotlib \cite{hunterMatplotlib2DGraphics2007} based on these descriptions because the Matplolib artist layer is analogous to the transforms.

\subsection{ What is a viz}?
Acquired codes of meaning

\section{Not all data are tables}

Tables, images (Lev), graphs (network X)

\end{document}

### 2020-12-08 draft/sections/math.tex
taus

For each field $c \in C$, the record function $r: C \rightarrow U_{\sigma}$ returns an object of type $\sigma(c) \in DT$. The set of all records $\Gamma(\sigma)$ is the set of all sections on $U_\sigma$ is $F$ over $E$. Spivak defines the $\tau$ mapping from an index of databases $K$ to records $\Gamma(\sigma)$ as $\tau: K \rightarrow \Gamma(\sigma)$. This is equivalent to $\tau: k \rightarrow E$ since $F = \Gamma(\sigma)$ and $F$ is the embedding in $E$ on which the records $r$ lie.

@@

%% pull out measurement
The fiber in figure~\ref{fig:temp} is the space of possible temperature values in degrees celsius, such that $F=[temp_{min}, temp_{max}]$ and each $r$ is a record with a single element $temp$. $F$ can be any number of dimensions, for example in figure~\ref{fig:temp_time} time is encoded as a second dimension. This means that the set of possible values $F$ is

@@

\begin{gather}
\tau(k) =(k, (temp, time))\\
temp \in [temp_{min}, temp_{max}], time \in [time_{min}, time_{max}])
\end{gather}

Since $\tau(k)=(k, r)$, $\sigma$ binds the $temp$ to a named temperature data field and a temperature data type and $time$ to a named time field and a time data type.

\subsubsection{Sheaf: Subset}
Something something this is how we get that subsets on E retain propertoes of transforms on E.
$\Gamma(E)$ is the space of all points in $F$ returned by $\sigma$; therefore the points being visualized in a streaming or animation example can be considered a subset that lives on base space $U \subset K$ with the same fiber $F$
\begin{equation}
    \begin{tikzcd}
        \iota^\ast E \arrow[r, hook] \arrow[d]                                                                       & E \arrow[d, bend right]                       \\
        U \arrow[r, "\iota" description, hook] \arrow[u, "\iota^\ast \sigma" description, bend right, shift right=2] & K \arrow[u, "\sigma" description, bend right]
    \end{tikzcd}    
\end{equation}
where $\iota^*E$ and $\iota^*\sigma$ are $E$ and $\sigma$ restricted to points $k \in U \subset K$.

@@

% preamble type intro to displays, examples - displays can include screen, 3D printer, sphereical display, H is the total space of the screen. 11
\begin{figure}[h]
    \includegraphics[width=.4\linewidth]{figures/sections/math/render.png}
    \caption{}
    \label{fig:render}
\end{figure}

The sets correspond to the sections on $\S$, which is the topology of the output of the artist $A$. The space $H$ is a total space representing the predisplay space, with a fiber of   which is dependent on the render space and a base space of $\S$:
\begin{equation}
    \begin{tikzcd}
        D \arrow[r, hook] & H \arrow[d, "\pi" description, bend right] \\
                                    & S \arrow[u, "\rho" description, bend right] 
    \end{tikzcd}
\end{equation}

We are primarily concerned with physical display space. For the renderers currently defined in Matplotlib, we can define the fiber as the $\mathbb{R}^{7}$ tuple
\begin{equation}
    \mathbb{R}^{7} = \{X, Y, Z, R, G, B, A\}
\end{equation}
where each tuple is representative of a pixel.

In the case of 2D screens, the predisplay space is a trivial fiber bundle $H=\mathbb{R}^{7}\times S$. As illustrated in figure~\ref{fig:render}, a region on the screen defined by the corners $(x_1, y_1)$ and $(x_2, y_2)$ maps into a region on a 2-simplex in $S$ defined by $(\alpha_1, \beta_1)$ and $(\alpha_2, \beta_2)$. The function on the simplex $f$ returns the (R, G, B, A) value for that $(\alpha, \beta)$ pair. For a region,

\begin{equation}
ab=23
\end{equation}

where the R,G,B,A values are derived from the how the data values are mapped to visual characteristics. The z component of the mapping to $\mathbb{R}^7$ is moved to the integration because this is a trivial space representing a 2D screen; $\rho$ varies depending on $H$.

\subsection{Artist}
%% include some words about motivation 
\begin{equation}
    A: \Gamma(E) \rightarrow \Gamma(H)
\end{equation}

\subsubsection{Screen to Data}
%%diagram: [data] -Tau->[RGVXYZ]
%%%            \ /\ / 
\begin{equation}
    \begin{tikzcd}
        E \arrow[d, "\pi" description] & H \arrow[d, "\pi" description]                                                 \\
        K \arrow[u, "\sigma" description, bend right] & S \arrow[l, "\xi" description] \arrow[u, "\rho" description, bend right]
    \end{tikzcd}
\end{equation}
The pullback $\xi$ on $S \rightarrow K$ means that the values in $E$ can be directly mapped to a simplex in $S$, which means there's a mapping from screen space back to the values. 
\begin{equation}
    \begin{tikzcd}
        \xi E \arrow[rr, "\tau" description] \arrow[rd, "\xi \sigma" description, bend right] &     & H \arrow[ld] \\
        & S \arrow[lu, bend right] &             
    \end{tikzcd}
\end{equation}

>>>>

For each field $c \in C$, the record function $r: C \rightarrow U_{\sigma}$ returns an object of type $\sigma(c) \in DT$. The set of all records $\Gamma(\sigma)$ is the set of all sections on $U_\sigma$. Spivak defines the $\tau$ mapping from an index of databases $K$ to records $\Gamma(\sigma)$ as $\tau: K \rightarrow \Gamma(\sigma)$. This is equivalent to $\tau: k \rightarrow E$ since $F = \Gamma(\sigma)$ and $F$ is the embedding in $E$ on which the records $r$ lie.

@@

%% pull out measurement
The fiber in figure~\ref{fig:temp} is the space of possible temperature values in degrees celsius, such that $F=[temp_{min}, temp_{max}]$ and is named \textrm{Temp}. In figure~\ref{fig:temp_time} \textrm{time} is encoded as a second dimension. This means that the set of possible values $F$ with $C=\{\textrm{Temp}, \textrm{Time}\}$:

@@

\begin{equation}
\tau(k) =(k, (r: \textrm{Temp}\rightarrow temp, r: \textrm{Time}\rightarrow time))\\
temp \in [temp_{min}, temp_{max}], time \in [time_{min}, time_{max}])
\end{equation}

Since $\tau(k)=(k, r)$, $temp$ is bound to a named data field and $sigma$ binds $temp$ to a temperature data type.

\subsubsection{Sheaf: Subset}
$\Gamma(E)$ is the space of all points in $F$ returned by $\sigma$; therefore the points being visualized in a streaming or animation example can be considered a subset that lives on base space $U \subset K$ with the same fiber $F$
\begin{equation}
    \begin{tikzcd}
        \iota^\ast E \arrow[r, hook] \arrow[d]                                                                       & E \arrow[d, bend right]                       \\
        U \arrow[r, "\iota" description, hook] \arrow[u, "\iota^\ast \sigma" description, bend right, shift right=2] & K \arrow[u, "\sigma" description, bend right]
    \end{tikzcd}    
\end{equation}
where $\iota^*E$ and $\iota^*\sigma$ are $E$ and $\sigma$ restricted to points $k \in U \subset K$.

@@

% preamble type intro to displays, examples - displays can include screen, 3D printer, sphereical display, H is the total space of the screen. 11

The sets correspond to the sections on $\S$, which is the topology of the output of the artist $A$. The space $H$ is a total space representing the predisplay space, with a fiber dependent on the render space and a base space of $\S$:
\begin{equation}
    \begin{tikzcd}
        D \arrow[r, hook] & H \arrow[d, "\pi" description, bend right] \\
                                    & S \arrow[u, "\rho" description, bend right] 
    \end{tikzcd}
\end{equation}

where $\rho: H \rightarrow S$ is mapping from a mathematical encoding of the image to pixels on the screen.

For a physical screen display, the fiber $D$ encodes position and color of the visual elements as $\mathbb{R}^{7}$:

\begin{equation}
    \mathbb{R}^{7} = \{X, Y, Z, R, G, B, A\}
\end{equation}
where the pixel $p \in H$ has the fields $\{x, y, z, r, g, b\}$. In the case of 2D screens, the predisplay space is a trivial fiber bundle $H=\mathbb{R}^{7}\times S$.

To draw an image, a region $p \in H$ is inverse mapped into a region $s \in S$
\begin{equation}
s = (\rho_{x}^{-1}(X(p)), \rho_{y}^{-1}(Y(p)))
\end{equation}

such that the rest of the fields in $\mathbb{R}^{7}$ are then integrated over $S$ to yield the remaining fields in $p$
\begin{align}
    R(p) &= \oint_s \rho_R(s)ds^{2}\\
    G(p) &= \oint_s \rho G(s)ds^{2}\\
    B(p) &= \oint_s \rho B(s)ds^{2}
\end{align}

Here we assume a single opaque 2D image such that the $z$ and $alpha$ fields can be omitted.

\begin{figure}[h]
    \includegraphics[width=.4\linewidth]{figures/sections/math/render.png}
    \caption{}
    \label{fig:render}
\end{figure}

As illustrated in figure~\ref{fig:render}, ....

\subsection{Artist}
The artist is a mapping from the $\Gamma(E)$ space of all datasets with a shared $F$ and $K$ to a pre-render space $H$ with display target properties $D$ and visualization structure $S$.

\begin{equation}
    A: \Gamma(E) \rightarrow \Gamma(H)
\end{equation}

\subsubsection{Screen to Data}
%%diagram: [data] -Tau->[RGVXYZ]
%%%            \ /\ / 
\begin{equation}
    \begin{tikzcd}
        E \arrow[d, "\pi" description] & H \arrow[d, "\pi" description]                                                 \\
        K \arrow[u, "\sigma" description, bend right] & S \arrow[l, "\xi" description] \arrow[u, "\rho" description, bend right]
    \end{tikzcd}
\end{equation}
The pullback $\xi$ on $S \rightarrow K$ means that $ r \in E$ can be mapped to from a simplex in $S$
\begin{equation}
    \begin{tikzcd}
        \xi E \arrow[rr, "\tau" description] \arrow[rd, "\xi \sigma" description, bend right] &                          & H \arrow[ld, "\rho^{-1}"] \\
                                                                                              & S \arrow[lu, bend right] &                          
        \end{tikzcd}
\end{equation}
